---
title: "Class 2 HW"
author: "Meghana Sannidhi"
date: "2026-09-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)

Typical_Employee_Survey_Data_1_ <-
read_excel("Typical_Employee_Survey_Data (1).xlsx",
col_names = FALSE)
```

# Class 2 Homework

## 1. Five-Point Summary and Boxplots
```{r}
ls()
summary(Typical_Employee_Survey_Data_1_)
str(Typical_Employee_Survey_Data_1_)
lapply(Typical_Employee_Survey_Data_1_, unique)
fivenum(Typical_Employee_Survey_Data_1_$...1)
fivenum(Typical_Employee_Survey_Data_1_$...5)

boxplot(Typical_Employee_Survey_Data_1_$...1,
        main = "Boxplot of Column 1",
        ylab = "Value")

boxplot(Typical_Employee_Survey_Data_1_$...5,
        main = "Boxplot of Column 5",
        ylab = "Value")
## 2. Convert Categorical Columns to Factors
```
Typical_Employee_Survey_Data_1_$...2 <- as.factor(Typical_Employee_Survey_Data_1_$...2)
Typical_Employee_Survey_Data_1_$...3 <- as.factor(Typical_Employee_Survey_Data_1_$...3)
Typical_Employee_Survey_Data_1_$...4 <- as.factor(Typical_Employee_Survey_Data_1_$...4)
Typical_Employee_Survey_Data_1_$...6 <- as.factor(Typical_Employee_Survey_Data_1_$...6)
Typical_Employee_Survey_Data_1_$...7 <- as.factor(Typical_Employee_Survey_Data_1_$...7)
Typical_Employee_Survey_Data_1_$...8 <- as.factor(Typical_Employee_Survey_Data_1_$...8)
Typical_Employee_Survey_Data_1_$...9 <- as.factor(Typical_Employee_Survey_Data_1_$...9)
Typical_Employee_Survey_Data_1_$...10 <- as.factor(Typical_Employee_Survey_Data_1_$...10)
Typical_Employee_Survey_Data_1_$...11 <- as.factor(Typical_Employee_Survey_Data_1_$...11)

str(Typical_Employee_Survey_Data_1_)

## 3. Visuals per column
```r
```{r}
hist(Typical_Employee_Survey_Data_1_$...1,
main = "Distribution of Column 1",
xlab = "Column 1")

hist(Typical_Employee_Survey_Data_1_$...5,
main = "Distribution of Column 5",
xlab = "Column 5")

barplot(table(Typical_Employee_Survey_Data_1_$...2),
main = "Distribution of Column 2",
xlab = "Category",
ylab = "Frequency")
barplot(table(Typical_Employee_Survey_Data_1_$...3),
main = "Distribution of Column 3",
xlab = "Category",
ylab = "Frequency")

barplot(table(Typical_Employee_Survey_Data_1_$...4),
main = "Distribution of Column 4",
xlab = "Category",
ylab = "Frequency")

barplot(table(Typical_Employee_Survey_Data_1_$...6),
main = "Distribution of Column 6",
xlab = "Category",
ylab = "Frequency")

barplot(table(Typical_Employee_Survey_Data_1_$...7),
main = "Distribution of Column 7",
xlab = "Category",
ylab = "Frequency")

barplot(table(Typical_Employee_Survey_Data_1_$...8),
main = "Distribution of Column 8",
xlab = "Category",
ylab = "Frequency")

barplot(table(Typical_Employee_Survey_Data_1_$...9),
main = "Distribution of Column 9",
xlab = "Category",
ylab = "Frequency")

barplot(table(Typical_Employee_Survey_Data_1_$...10),
main = "Distribution of Column 10",
xlab = "Category",
ylab = "Frequency")

barplot(table(Typical_Employee_Survey_Data_1_$...11),
main = "Distribution of Column 11",
xlab = "Category",
ylab = "Frequency")
```
## 4. Scatterplots show relationships
```{r}
plot(Typical_Employee_Survey_Data_1_$...1,
Typical_Employee_Survey_Data_1_$...5,
main = "Relationship Between Column 1 and Column 5",
xlab = "Column 1",
ylab = "Column 5",
pch = 19)
```





